---
title: "Orange Tree Predictor"
author: "Christian Thyssen"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The Data

Our application is based on the data frame 'Orange' from the package 'datasets'.

```{r, echo=TRUE}
head(Orange)
```

It contains the age and the circumference of a number of orange trees.

## The Application - Overview

Our application has the following features:

- The app visualizes the data.
- The user can configure the plot.
- The app builds and visualizes a regression model for predicting the circumference of an orange tree from its age.
- The user can tune the model.
- The user can use the model to predict the circumference of an orange tree.
- The app includes a help page.

## The Model

We decided to use a basis spline to predict the circumference of an orange tree from its age.

Our model has two tuning parameters:

- The parameter 'number of internal knots' controls into how many pieces the range is decomposed (from 0 to 5). For n knots there are n+1 pieces.
- The parameter 'degree' controls the degree of the piecewise polynomials (from 1 to 5).

With the default settings (number of knots = 0 and degree = 1), the model is a linear model.

## The Application - Details

The following plot shows an example from the app:

```{r, warning=FALSE, message=FALSE}
library(splines); library(tidyverse)
model <- lm(circumference ~ bs(age, df = 1 + 3, degree = 3), data = Orange)
fun <- function(x) {
    newdata <- data.frame(age = x)
    predict(model, newdata = newdata)
}
Orange %>%
    ggplot(aes(x = age, y = circumference, colour = Tree)) +
    labs(title = "Growth of Orange Trees") +
    geom_function(fun = fun, colour = "red") +
    geom_point(size = 3) +
    geom_point(aes(x = 900, y = fun(900)), colour = "red", size = 3)
```
